\begin{thebibliography}{100}

\bibitem{firefox} The Firefox web browser.
  Located at: https://www.mozilla.org/en-GB/firefox/new/ [Accessed 14/01/2016]

\bibitem{browserstats} Browser Statistics and Trends.
  Located at: http://www.w3schools.com/browsers/browsers\_stats.asp [Accessed 14/01/2016]

\bibitem{openoffice} The OpenOffice word processing suite.
  Located at: https://www.openoffice.org [Accessed 14/01/2016]

\bibitem{osalternatives} Comparing Microsoft Office to Open Source Alternatives,
  Quinn, L. S and Andrei, K. H.,
  2013.
  Located at: http://www.idealware.org/articles/comparing-microsoft-office-open-source-alternatives [Accessed 14/01/2016]

\bibitem{websurvey} Web Survey Data,
  2014.
  Located at: https://secure1.securityspace.com/s\_survey/data/201402/index.html [Accessed 14/01/2016]

\bibitem{osusage} Usage of operating systems for websites.
  Located at: http://w3techs.com/technologies/overview/operating\_system/all [Accessed 14/01/2016]

\bibitem{commonsoftware} Guide to Software on Campus Computers,
  Catholic University of America.
  Located at: http://computing.cua.edu/support/common-software-list.cfm  [Accessed 14/01/2016]

\bibitem{commonsoftware2} What applications are most commonly used?
  CERN IT Technology Department.
  Located at: http://information-technology.web.cern.ch/book/desktop-os-comparison/what-applications-are-most-commonly-used [Accessed 14/01/2016]

\bibitem{buildessential} Informational list of build-essential packages.
  Located at: https://packages.debian.org/wheezy/build-essential [Accessed 16/01/2016]

\bibitem{fsf} The Free Software Foundation.
  Located at: http://fsf.org [Accessed 15/01/2016]

\bibitem{ibmlinux} IBM and the Linux Community.
	Located at: http://www-03.ibm.com/linux/community.html [Accessed 17/01/2016]

\bibitem{mozillacrisis} Personality and Change Inflamed Mozilla Crisis,
    Q.Hardy and N. Bilton,
    2014.
	Located at: http://www.nytimes.com/2014/04/05/technology/personality-and-change-inflamed-crisis-at-mozilla.html [Accessed 17/01/2016]

\bibitem{osihistory} The History of the OSI,
  The Open Source Initiative.
  Located at: http://opensource.org/history [Accessed 15/01/2016]

\bibitem{copyleft} What is copyleft?
  The GNU Foundation.
  Located at: https://www.gnu.org/copyleft/ [Accessed 16/01/2016]

\bibitem{cc} The Creative Commons Licenses.
  Located at: https://creativecommons.org/ [Accessed 15/01/2016]

\bibitem{ccdir} Content Directories: list of organizations and projects powered with Creative Commons licenses.
  Located at: https://wiki.creativecommons.org/wiki/Content\_Directories [Accessed 15/01/2016]

\bibitem{linuxkernel} Breakdown of code in the Linux Kernel,
  Black Duck Software.
  Located at: https://www.openhub.net/p/linux/analyses/latest/languages\_summary [Accessed 15/01/2016]

\bibitem{historyos} A brief history of open source software,
  Gonzalez-Barahona J. M.,
  2000.
  Located at: http://eu.conecta.it/paper/brief\_history\_open\_source.html [Retrieved 21/01/2016]

\bibitem{faif} Free as in Freedom,
  Williams, S.,
  2002.
  O\'Reilly Media.

\bibitem{stallmanusenet} UseNet post announcing a "new UNIX implementation",
  Stallman, R.,
  1983.
  Located at: https://groups.google.com/forum/\#!msg/net.unix-wizards/8twfRPM79u0/1xlglzrWrU0J [Accessed 12/01/2016]

\bibitem{osi} The Open Source Initiative Homepage.
  Located at: https://opensource.org/ [Retrieved 21/01/2016]

\bibitem{osireview} The License Review Process,
  The Open Source Initiative.
  Located at: https://opensource.org/approval [Accessed 16/01/2016]

\bibitem{gpl} The GNU Public License.
  Located at: https://opensource.org/licenses/gpl-license [Accessed 16/01/2016]

\bibitem{surveyosl} Surveying Open Source Licenses,
  Kerrisk, M.,
  2013.
  Located at: https://lwn.net/Articles/547400/ [Accessed 17/01/2016]

\bibitem{viral} Speech Transcript\: The New York University Stern School of Business,
    Craig Mundie.
    Located at: http://news.microsoft.com/speeches/speech-transcript-craig-mundie-the-new-york-university-stern-school-of-business/ [Retrieved 16/01/2016]

\bibitem{kempcurrentdev} Current developments in Open Source Software,
    Kemp, R.,
    Computer Law and Security Review,
    2009.

\bibitem{tivo} The TiVo Homepage.
  Located at: https://www.tivo.com/ [Retrieved 21/01/2016]

\bibitem{bsd} The BSD License.
	Located at: https://opensource.org/licenses/BSD-3-Clause [Retrieved 17/01/2016]

\bibitem{mit} The MIT License.
	Located at: https://opensource.org/licenses/MIT [Retrieved 17/01/2016]

\bibitem{apache} The Apache License.
	Located at: http://www.apache.org/licenses/LICENSE-2.0 [Retrieved 17/01/2016]

\bibitem{oslicenseusage} Top 20 Open Source Licenses,
	Black Duck Software.
	Located at: https://www.blackducksoftware.com/resources/data/top-20-open-source-licenses [Accessed 18/01/2016]
    
\bibitem{cio} Enterprise Developers Programming Speed? Check. Time to Fix Bugs? Not So Fast,
  Schindler, E.,
  2007.
  Located at: http://www.cio.com/article/2374313/developer/enterprise-developers-programming-speed--check--time-to-fix-bugs--not-so-fast-.html [ Retrieved 21/01/2016]

\bibitem{nginx} The Nginx Homepage
  Located at: https://www.nginx.com/ [Retrieved 17/01/2016]

\bibitem{sendmail} The Sendmail Homepage
  Located at: http://www.sendmail.com/sm/open\_source/ [Retrieved 17/01/2016]

\bibitem{docker} The Docker Homepage.
  Located at: https://www.docker.com/ [Retrieved 17/01/2016]

\bibitem{catb} The Cathedral and the Bazaar
  Raymond, E. S.
  O'Reilly Media,
  1999.

\bibitem{heartbleed} The Heartbleed OpenSSL Vulnerability.
  Located at: http://heartbleed.com/ [Retrieved 17/01/2016]

\bibitem{secureoss} Open Source Security: A Look at the Security Benefits of Source Code Access,
  TruSecure and Red Hat,
  2001.
  Located at: http://www.redhat.com/whitepapers/services/Open\_Source\_Security5.pdf [Retrieved 17/01/2016]

\bibitem{feedfish} Feed the Fish,
  Blevins, D.,
  2013.
  Located at: tomitribe.com/blog/2013/11/feed-the-fish/ [Retrieved 16/01/2016]

\bibitem{stackoverflow} The StackOverflow Homepage.
  Located at: https://stackoverflow.com [Retrieved 17/01/2016]

\bibitem{whyemacs} Why Emacs?
  Batsov, B.,
  2011.
  Located at: http://batsov.com/articles/2011/11/19/why-emacs/ [Retrieved 17/01/2016]
  
\bibitem{wordopenoffice} Question: Why Microsoft Word or LibreOffice/OpenOffice? on Research Gate,
2012.
Located at: https://www.researchgate.net/post/Why\_Microsoft\_Word\_or\_LibreOffice\_OpenOffice [Retreived 21/01/2016]

\bibitem{mysql} The MySQL Homepage.
  Located at: http://www.mysql.com/ [Retrieved 18/01/2016]

\bibitem{monty} The Business of Open Source Software: A Primer,
    Widenius, M. and Nyman, L.,
    Technology Innovation Management Review,
    2014.

\bibitem{magiclantern} The Magic Lantern Homepage.
  Located at: http://magiclantern.wikia.com/wiki/Magic\_Lantern\_Firmware\_Wiki [Retrieved 17/01/2016]

\bibitem{howopen} How open is open enough? Melding proprietary and open source platform strategies,
  West, J.,
  Research Policy 32,
  pp 1259 - 1285,
  2003.

\bibitem{lufthansa} Lufthansa close to deal with IBM for IT infrastructure unit,
  V. Bryan, H. ten Wolde and E. Auchard.
  Located at: http://www.reuters.com/article/lufthansa-ma-it-idUSL6N0SH14320141022 [Retrieved 16/01/2016]

\bibitem{redhat} The Red Hat Homepage.
  Located at: http://www.redhat.com/ [Retrieved 15/01/2016]

\bibitem{distros} Top Ten Linux Distributions.
  Located at: http://distrowatch.com/dwres.php?resource=major [Retrieved 17/01/2016]

\bibitem{redhatmoney} Red Hat Reports Fourth Quarter and Fiscal Year 2015 Results,
  March 2015.
  Located at: http://investors.redhat.com/releasedetail.cfm?ReleaseID=903376 [Retrieved 17/01/2016]
  
\bibitem{dontredhat}
Open source startups: Don't try to be Red Hat,
Phipps, S.,
2014.
Located at: http://www.infoworld.com/article/2610426/open-source-software/open-source-startups--don-t-try-to-be-red-hat.html [Retrieved 21/01/2016]

\bibitem{advbudget} Advertising is a key strategy for Coca-Colas growth,
  Bailey, S.
  Located at:  http://finance.yahoo.com/news/advertising-key-strategy-coca-cola-130112295.html [Retrieved 17/01/2016]

\bibitem{bootstrap} The Twitter Bootstrap Homepage.
  Located at:  http://getbootstrap.com/ [Retrieved 17/01/2016]

\bibitem{bootstrappress} Bootstrap from Twitter, 
  Twitter, Inc.
  Located at: https://blog.twitter.com/2011/bootstrap-from-twitter (Online). [Retrieved 17/01/2016]

\bibitem{bootstrapblog} Bootstrap Project Blog.
  Located at: http://blog.getbootstrap.com/ [Retrieved 17/01/2016]

\bibitem{bootstrapmoney} How does Twitter make money from open sourcing Bootstrap?
  M. Otto.
  Located at: https://www.quora.com/How-does-Twitter-make-money-from-open-sourcing-Bootstrap [Retrieved 18/01/2016]

\bibitem{wikimedia} The Wikimedia Foundation.
  Located at: https://wikimediafoundation.org [Retrieved 17/01/2016]

\bibitem{bitcoin} The Bitcoin Foundation.
  Located at: https://bitcoinfoundation.org/ [Retrieved 17/01/2016]

\bibitem{apple} The Apple Homepage.
  Located at: http://apple.com [Retrieved 17/01/2016]
  
\bibitem {apsl}
  The Apple Open Source License.
  Located at: http://www.opensource.apple.com/license/apsl/  [Retrieved 21/01/2016]

\bibitem{appleos} Open Source at Apple.
  Located at: http://www.apple.com/opensource/ [Retrieved 17/01/2016]

\bibitem{webkitfork} Google going its own way, forking WebKit rendering engine,
    Ars Technica.
    Located at: http://arstechnica.com/information-technology/2013/04/google-going-its-own-way-forking-webkit-rendering-engine/ [Retrieved 19/01/2016]

\bibitem{tesla} Tesla Motors.
  Located at: http://teslamotors.com [Retrieved 17/01/2016]

\bibitem{teslapatents} All our patents are belong to you,
  Tesla Motors.
  Located at: https://www.teslamotors.com/blog/all-our-patent-are-belong-you [Retrieved 17/01/2016]

\bibitem{teslasafety} Tesla Model S achieves best safety rating of any car ever tested.
	Located at: https://www.teslamotors.com/en\_GB/blog/tesla-model-s-achieves-best-safety-rating-any-car-ever-tested [Retrieved 19/01/2016]

\bibitem {spacex}
  The SpaceX Homepage.
  Located at: http://www.spacex.com/  [Retrieved 21/01/2016]
  
\bibitem {solarcity}
  The Solar City Homepage.
  Located at: http://www.solarcity.com/  [Retrieved 21/01/2016]

\bibitem{elonmusk} Elon Musk: The World's Raddest Man,
    Wait But Why,
    2015.
	Located at: http://waitbutwhy.com/2015/05/elon-musk-the-worlds-raddest-man.html [Accessed 18/01/2016]

\bibitem{googlebenefits} Google Employee Benefits,
  Google, Inc.
  Located at: http://www.google.co.uk/about/careers/lifeatgoogle/benefits/ [Retrieved 17/01/2016]

\bibitem{appleenvironment} Environmental Responsibility,
  Apple, Inc.
  Located at: http://www.apple.com/uk/environment/ [Retrieved 17/01/2016]

\bibitem{dontbeevil} Google Code of Conduct,
  Google, Inc.
  Located at: https://investor.google.com/corporate/google-code-of-conduct.html [Retrieved 17/01/2016]

\bibitem{github} The GitHub Homepage.
  Located at: https://www.github.com/  [Retrieved 21/01/2016]

\bibitem{editorusage} The Stack Overflow Developer Survey,
    2015.
	Located at: http://stackoverflow.com/research/developer-survey-2015 [Accessed 16/01/2016]

\bibitem{bootstrapcontrib} The contributors for Twitter Bootstrap on GitHub.
	Located at: https://github.com/orgs/twbs/people [Accessed 17/01/2016]

\bibitem{gigafactory} Planned 2020 Gigafactory Production Exceeds 2013 Global Production.
	Located at: https://www.teslamotors.com/sites/default/files/blog\_attachments/gigafactory.pdf [Retrived 19/01/2016]

\bibitem{google} The Google Homepage.
    Located at: https://google.com [Accessed 18/01/2016]

\bibitem{android} The Android Homepage.
    Located at: https://android.com [Accessed 18/01/2016]

\bibitem{chromium} The Chromium Project Homepage.
    Located at: https://chromium.org [Accessed 18/01/2016]

\bibitem{linearbookscanner} The Linear Book Scanner,
  Google, Inc.
  Located at: https://code.google.com/p/linear-book-scanner/ [Retrieved 15/01/2016]


\end{thebibliography}

%%%%% CLEAR DOUBLE PAGE!
\newpage{\pagestyle{empty}\cleardoublepage}